# Blueprint simulator

``` {r Setup, echo = FALSE, message = FALSE}

library(readxl)
library(plyr)
library(dplyr)
library(ggplot2)
library(furrr)
library(stringr)
library(future)

# Custom variables
path_blueprint_xlsx <- "G:/My Drive/Gaming/Oxygen Not Included/blueprint_supply_closet.xlsx"
path_simdata        <- "../Data/simdata_blueprint_n3.RData"
path_simulation_R   <- "../R/simulate_blueprints.R"
n_sim               <- 1000   # Default: 1000
use_seed            <- as.integer(Sys.Date())
force               <- FALSE # Default FALSE; Whether to force simulation or use simdata if it exists
```

## Description

A script that simulates how long it takes to gather all blueprints using the old system with 3 weekly drops.

``` {r Loading data, echo = FALSE, message = FALSE}
blueprints <- read_xlsx(path = path_blueprint_xlsx,
                        sheet = "Blueprints",
                        col_names = c("id", "primary", "secondary", "item_name", "tier", "owned"),
                        skip = 1) |>
  mutate(across(c(primary, secondary, tier), as.factor),
         owned = as.logical(ifelse(owned == "Yes", 1, 0)))

blueprint_data <- read_xlsx(path = path_blueprint_xlsx,
                            sheet = "Data",
                            col_names = c("tier", "cost", "n_items", "prob", "prob_item"),
                            skip = 1) |>
  mutate(tier = as.factor(tier))

n_total <- sum(blueprint_data$n_items)

source(file = path_simulation_R)
```

By default the simulation works based on previously observed probabilities which are:

* Common:   `r signif(filter(blueprint_data, tier == "Common")$prob, digits = 2)`
* Decent:   `r signif(filter(blueprint_data, tier == "Decent")$prob, digits = 2)`
* Nifty:    `r signif(filter(blueprint_data, tier == "Nifty")$prob, digits = 2)`
* Splendid: `r signif(filter(blueprint_data, tier == "Splendid")$prob, digits = 2)`

Script works as the following:

1. Draws a random blueprint based on tier probabilities (also called an attempt)
2. If that blueprint has already been collected, then recycle it, if not keep it inventory
3. **Spender strategy:**

    * Save up until get `r max(blueprint_data$cost)` filaments (in order to buy any tier)
    * Buy a missing blueprint at random (any tier)
    * Reduce the amount of filament according to cost
    * Continue to buy blueprints if filaments are still greater than `r max(blueprint_data$cost)`.
    * If only one missing blueprint left then buy it
  
4.  **Hoarder strategy:**

    * Saves up all filaments from each recycling
    * Calculate the total filament cost of remaining missing blueprints
    * If saved up filaments is equal or greater then the remaining cost for missing blueprints, then stop as you now can buy the remaining blueprints
  
5. Print out the number of attempts, time and filaments required (if saving up).

## Simulation conditions

A total of `r n_total` blueprints ready to be claimed.

The number of blueprints claimable per week is 3.

The average time in hours it takes for claiming all blueprints in a week is estimated to be around 1 hour and 50 minutes.

The recycled yield is calculated as getting 25% back when recycled.

``` {r Run simulations, echo = FALSE, message = FALSE}

if (force | !file.exists(path_simdata)) {
  
  # Simulate using parallel processing
  plan(multisession, workers = availableCores() - 2)
  
  spender <- future_map(seq_len(n_sim),
                        .options = furrr_options(seed = use_seed),
                        ~ simulate_blueprints(use_owned = FALSE,
                                              hoard = FALSE,
                                              starting_filament = 0,
                                              n_week = 3,
                                              time_week = 1.83,
                                              recycle_yield = 0.25))
  
  hoarder <- future_map(seq_len(n_sim),
                        .options = furrr_options(seed = use_seed),
                        ~ simulate_blueprints(use_owned = FALSE,
                                              hoard = TRUE,
                                              starting_filament = 0,
                                              n_week = 3,
                                              time_week = 1.83,
                                              recycle_yield = 0.25))
  
  plan(sequential)
  
  sim_data <- bind_rows(
    mutate(approach = "Spender", do.call(rbind, spender)),
    mutate(approach = "Hoarder", do.call(rbind, hoarder))
  ) |>
    relocate(approach, .before = total_attempts)
  
  sim_single <- simulate_blueprints(use_owned = FALSE, hoard = FALSE, detailed = TRUE)$detailed
  
  save(sim_data, sim_single, file = path_simdata)
} else {
  load(file = path_simdata)
}

sim_spender <- filter(sim_data, approach == "Spender")
sim_hoarder <- filter(sim_data, approach == "Hoarder")

```

## Results

Ran using `r n_sim` simulations per approach.

### Spender strategy

Average hours required: `r round(mean(sim_spender$hours), digits = 0)`

Average weeks required: `r round(mean(sim_spender$weeks), digits = 0)`

Average years required: `r round(mean(sim_spender$years), digits = 1)`

Average number of recycled blueprints: `r round(mean(sim_spender$recycled), digits = 0)`

### Hoarder strategy

Average hours required: `r round(mean(sim_hoarder$hours), digits = 0)`

Average weeks required: `r round(mean(sim_hoarder$weeks), digits = 0)`

Average years required: `r round(mean(sim_hoarder$years), digits = 1)`

Average number of recycled blueprints: `r round(mean(sim_hoarder$recycled), digits = 0)`

Average filament required for tipping point: `r round(mean(sim_hoarder$filament_tip / 1000, na.rm = TRUE), digits = 0)`K

``` {r Data-Vis, echo = FALSE, message = FALSE}

title_hist <- "Distribution of time to collect all blueprints (3 per week)"

ggplot(sim_data, aes(x = hours, fill = approach)) +
  geom_histogram(binwidth = 10, alpha = 0.7, color = "black", position = "dodge") +
  labs(
    title = title_hist,
    x = "Time (Hours)",
    y = "Frequency"
  ) +
  scale_fill_discrete(name = "Approach") +
  scale_x_continuous(breaks = seq(from = round_any(min(sim_data$hours), 50, f = floor), to = round_any(max(sim_data$hours), 50, f = ceiling), by = 50)) +
  theme_minimal()

ggplot(sim_data, aes(x = approach, y = hours, fill = approach)) +
  geom_boxplot(show.legend = FALSE, color = "black") +
  labs(title = title_hist,
       x = "Approach",
       y = "Hours") +
  scale_y_continuous(breaks = seq(from = round_any(min(sim_data$hours), 50, f = floor), to = round_any(max(sim_data$hours), 50, f = ceiling), by = 50))

ggplot(sim_data, aes(x = weeks, fill = approach)) +
  geom_histogram(binwidth = 2, alpha = 0.7, color = "black", position = "dodge") +
  labs(
    title = title_hist,
    x = "Time (Weeks)",
    y = "Frequency"
  ) +
  scale_fill_discrete(name = "Approach") +
  scale_x_continuous(breaks = seq(from = round_any(min(sim_data$weeks), 10, f = floor), to = round_any(max(sim_data$weeks), 10, f = ceiling), by = 10)) +
  theme_minimal()

ggplot(sim_data, aes(x = approach, y = weeks, fill = approach)) +
  geom_boxplot(show.legend = FALSE, color = "black") +
  labs(title = title_hist,
       x = "Approach",
       y = "Weeks") +
  scale_fill_discrete(name = "Approach") +
  scale_y_continuous(breaks = seq(from = round_any(min(sim_data$weeks), 10, f = floor), to = round_any(max(sim_data$weeks), 10, f = ceiling), by = 10))

ggplot(sim_data, aes(x = years, fill = approach)) +
  geom_histogram(binwidth = 0.1, alpha = 0.7, color = "black", position = "dodge") +
  labs(
    title = title_hist,
    x = "Time (Years)",
    y = "Frequency"
  ) +
  scale_fill_discrete(name = "Approach") +
  scale_x_continuous(breaks = seq(from = round_any(min(sim_data$years), 0.5, f = floor), to = round_any(max(sim_data$years), 0.5, f = ceiling), by = 0.5)) +
  theme_minimal()

ggplot(sim_data, aes(x = approach, y = years, fill = approach)) +
  geom_boxplot(show.legend = FALSE, color = "black") +
  labs(title = title_hist,
       x = "Approach",
       y = "Years") +
  scale_fill_discrete(name = "Approach") +
  scale_y_continuous(breaks = seq(from = round_any(min(sim_data$years), 0.5, f = floor), to = round_any(max(sim_data$years), 0.5, f = ceiling), by = 0.5))

sim_data |>
  filter(approach == "Hoarder") |>
  ggplot(aes(x = filament_tip / 1000)) +
  geom_histogram(binwidth = 1, alpha = 0.7, color = "black", fill = "cornflowerblue") +
  labs(
    title = "Distribution of filament tipping point",
    subtitle = "No owned blueprints, using a hoarder strategy",
    x = "Filament (per 1000)",
    y = "Frequency"
  ) +
  theme_minimal()

sim_data |>
  filter(approach == "Hoarder") |>
  ggplot(aes(x = approach, y = filament_tip / 1000)) +
  geom_boxplot(show.legend = FALSE, color = "black", fill = "cornflowerblue") +
  labs(title = "Boxplot distribution of filament tipping point",
       x = "Approach",
       y = "Filament (per 1000)")

```